---
title: "meas_report_month_CreateDatasets"
author: "Robesyn"
date: "5 January 2018"
output: html_document
---

```{r}
library(knitr)
library(rmarkdown)
library(anytime)
library(tidyverse)
library(lubridate)
library(stringr)
```

```{r}
meas <- read_csv("../2 data/1 raw data/export_20180726/2018-07-26_WHO_TESSyExport_Measles_case_based_data.csv")   
```

```{r}
glimpse(meas)        # includes discarded
```

```{r}
meas <- meas %>% select(
                              "NationalIdentifier",              
                              "ReportingCountry"   ,          
                              "DateUsedForStatisticsISO"  ,   
                              "DateUsedForStatisticsYear"  ,  
                              "DateUsedForStatisticsMonth",
                              #"DateOfOnsetISOdate",
                              "DateOfOnsetYear",
                              #"DateOfOnsetMonth",
                              "PlaceOfNotification",
                              "PlaceOfResidence",
                              "Age",
                              "AgeMonth",
                              "Gender",
                              "Classification",
                              "ClinicalCriteria",
                              "LaboratoryResult",
                              "EpiLinked",
                              "VaccStatus",
                              "Hospitalisation",
                              "Complications",
                              "Outcome",
                              "CauseOfDeathText",
                              "Imported",
                              "ProbableCountryOfInfection",
                              "ClusterSetting",
                              "ClusterIdentification",
                              "ResultIgM",                    
                              "ResultIgG",
                              "SpecimenSero",
                              "ResultVirDetect",
                              "SpecimenVirDetect", 
                              "Genotype",
                              "isNotification",
                              "isLaboratory"
                              )
```

```{r}
names(meas) <- tolower(names(meas))  
```

# Labelling unknown as missing
```{r}
meas$gender[meas$gender == "UNK"] <- NA
meas$outcome[meas$outcome == "UNK"] <- NA
meas$hospitalisation[meas$hospitalisation == "UNK"] <- NA  
meas$complications[meas$complications == "UNK"] <- NA  
# meas$classification[meas$classification == "UNK"] <- NA    # classification not made NA here, to avoid these rows get dropped by filter furter  
meas$vaccstatus[meas$vaccstatus == "UNK"] <- NA  
meas$imported[meas$imported == "UNK"] <- NA  
meas$clustersetting[meas$clustersetting == "UNK"] <- NA 
```

########################################
# replace countrycodes by countrynames #
########################################

```{r replace countrycodes by countrynames}
geocodes <- read_csv("../2 data/3 emma geocodes/geocodes.csv")
(geocodes <- geocodes %>% select(GEO_NAME, GEO_ID, C_16))   # 38999 rows, first 255 are countrycodes
```

```{r}
geocodes <- geocodes %>% filter(C_16==".All countries") %>% select(-C_16)
(geocodes <- rename(geocodes, geoname=GEO_NAME, geoid=GEO_ID))
```

```{r}
# create same varname as in case dataset to allow left_join
geocodes <- geocodes %>% mutate("reportingcountry"=geoid)
(geocodes_reportingcountry <- geocodes %>% select(geoname, reportingcountry))
```

````{r}
# left_join
meas <- left_join(meas, geocodes_reportingcountry, by="reportingcountry")
meas <- rename(meas, reportingcountryname=geoname)
meas %>% select(reportingcountry, reportingcountryname)
sum(is.na(meas$reportingcountryname))
```

```{r}
# overwrite original countrycodes with countrynames, and delete temporary vars
meas$reportingcountry <- meas$reportingcountryname
(meas <- select(meas, -reportingcountryname))
```

################
# set datatype #
################

```{r}
meas$reportingcountry <- as.factor(meas$reportingcountry)
meas$outcome <- as.factor(meas$outcome)

meas$dateusedforstatisticsiso <- as_date(meas$dateusedforstatisticsiso)

meas$placeofnotification  <- as.factor(meas$placeofnotification)
meas$placeofresidence  <- as.factor(meas$placeofresidence)
meas$gender <- as.factor(meas$gender)
meas$classification <- as.factor(meas$classification)
meas$clinicalcriteria <- as.factor(meas$clinicalcriteria)
meas$laboratoryresult <- as.factor(meas$laboratoryresult)
meas$epilinked <- as.factor(meas$epilinked)
meas$hospitalisation <- as.factor(meas$hospitalisation)
meas$vaccstatus <- as.factor(meas$vaccstatus)
meas$complications <- as.factor(meas$complications)
meas$outcome <- as.factor(meas$outcome)
meas$causeofdeathtext <- as.factor(meas$causeofdeathtext)
meas$imported <- as.factor(meas$imported)
meas$probablecountryofinfection <- as.factor(meas$probablecountryofinfection)
meas$clustersetting <- as.factor(meas$clustersetting)
meas$clusteridentification <- as.factor(meas$clusteridentification)
meas$clusteridentification <- as.factor(meas$clusteridentification)
meas$resultigm  <- as.factor(meas$resultigm)
meas$resultigg  <- as.factor(meas$resultigg)
meas$specimensero  <- as.factor(meas$specimensero)
meas$resultvirdetect <- as.factor(meas$resultvirdetect)
meas$specimenvirdetect  <- as.factor(meas$specimenvirdetect )
meas$genotype <- as.factor(meas$genotype)
meas$isnotification <- as.factor(meas$isnotification)
meas$islaboratory <- as.factor(meas$islaboratory)
```

# Exclude incompletely reported months
```{r}
meas <- meas %>% filter(!(reportingcountry == "Czech Republic" & dateusedforstatisticsyear == 2018  & dateusedforstatisticsmonth == 5)) 
```

# setting report period
```{r}
# Adapt code to avoid problem in month January (report in January about data November)
# today()

# thisyear <- year(today())
# lastyear <- thisyear - 1

# thismonth <- month(today())
# lastmonth <- thismonth - 1  #  to be split in months 2 to 12 and month 1

# thisyear
# thismonth
# lastyear
# lastmonth

# EASIER set explicit year and month
yearofreport <- 2018
monthofreport <- 6  # month of cases we report about (not month that report is made)
```

# DATASET 1: month
```{r}
# discarded
(meas_month_discarded <- meas %>% filter(dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth == monthofreport & classification == "DISCARDED") )  

# not discarded
(meas_month_notdiscarded <- meas %>% filter(dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth == monthofreport & classification != "DISCARDED")  )   
meas_month_notdiscarded 
```

# DATASET 2: last 12 months
```{r}
# discarded
meas_last12months_discarded <- meas %>% filter(
               (dateusedforstatisticsyear == yearofreport-1 & dateusedforstatisticsmonth > monthofreport & classification == "DISCARDED") |
               (dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth <= monthofreport & classification == "DISCARDED"))  
meas_last12months_discarded     

# not discarded
meas_last12months_notdiscarded <- meas %>% filter(
               (dateusedforstatisticsyear == yearofreport-1 & dateusedforstatisticsmonth > monthofreport & classification != "DISCARDED") |
               (dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth <= monthofreport & classification != "DISCARDED"))  
meas_last12months_notdiscarded    
```

# DATASET 3: cases calendaryear 2018
```{r}
# discarded
meas_calendaryear_discarded <- meas %>% filter(dateusedforstatisticsyear == yearofreport & classification == "DISCARDED")  
meas_calendaryear_discarded

# not discarded
meas_calendaryear_notdiscarded <- meas %>% filter(dateusedforstatisticsyear == yearofreport & classification != "DISCARDED") 
meas_calendaryear_notdiscarded
```

# DATASET 4 (4a, 4b, 4c): aggregated data 
Rows are case counts, not individual cases  -->  sum cases instead of count rows (variable 'numberofcases')

# read
```{r}
measagg <- read_csv("../2 data/1 raw data/export_20180726/2018-07-26_WHO_TESSyExport_Measles_vaggregated_data.csv")
names(measagg) <- tolower(names(measagg))  
```

```{r}
(measagg <- measagg %>% select(numberofcases, classification, dateusedforstatisticsyear, dateusedforstatisticsmonth, everything())  )
glimpse(measagg)  # cave incl discarded
```

# Labelling unknown as missing
```{r}
measagg$gender[measagg$gender == "UNK"] <- NA
measagg$vaccstatus[measagg$vaccstatus == "UNK"] <- NA  
```

```{r}
measagg <- measagg %>% mutate(reportingcountry = ifelse(reportingcountry == "BE", "Belgium", "Country to check"))
```

```{r}
table(measagg$reportingcountry)
table(measagg$classification)   # cave are rows of aggregated data, not cases! (see variable numberofcases)
```

## DATASET 4a: month agg 
```{r}
# discarded
(meas_agg_month_discarded <- measagg %>% filter(dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth == monthofreport & classification == "DISCARDED") )

# not discarded
(meas_agg_month_notdiscarded <- measagg %>% filter(dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth == monthofreport & classification != "DISCARDED") )
```

## DATASET 4b: last 12 months agg
```{r}
# discarded
(meas_agg_last12months_discarded <- measagg %>% filter(
               (dateusedforstatisticsyear == yearofreport-1 & dateusedforstatisticsmonth > monthofreport & classification == "DISCARDED") |
               (dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth <= monthofreport & classification == "DISCARDED"))  )

# not discarded
(meas_agg_last12months_notdiscarded <- measagg %>% filter(
               (dateusedforstatisticsyear == yearofreport-1 & dateusedforstatisticsmonth > monthofreport & classification != "DISCARDED") |
               (dateusedforstatisticsyear == yearofreport & dateusedforstatisticsmonth <= monthofreport & classification != "DISCARDED"))  )
```

## DATASET 4c: cases 2018 agg
```{r}
# discarded
(meas_agg_calendaryear_discarded <- measagg %>% filter(dateusedforstatisticsyear == yearofreport & classification == "DISCARDED")  )

# not discarded
(meas_agg_calendaryear_notdiscarded <- measagg %>% filter(dateusedforstatisticsyear == yearofreport & classification != "DISCARDED")  )
```

# DATASET 5: Combined dataset (join case based and agg)

1/ uncount() agg
```{r}
meas_agg_month_notdiscarded 
meas_agg_month_notdiscarded %>% summarize(total = sum(numberofcases))

(meas_disaggregated_month_notdiscarded <- meas_agg_month_notdiscarded %>% uncount(numberofcases, .id = "id_rank_from_agg") %>% select(id_rank_from_agg, everything())  )  
(meas_disaggregated_last12months_notdiscarded <- meas_agg_last12months_notdiscarded %>% uncount(numberofcases, .id = "id_rank_from_agg") %>% select(id_rank_from_agg, everything())  )
(meas_disaggregated_calendaryear_notdiscarded <- meas_agg_calendaryear_notdiscarded %>% uncount(numberofcases, .id = "id_rank_from_agg") %>% select(id_rank_from_agg, everything())  )
```

2/ prepare data class for join
```{r}
# fix incompatibilities to allow join
glimpse(meas_agg_month_notdiscarded)

meas_disaggregated_month_notdiscarded$dateusedforstatisticsiso <- anydate(meas_disaggregated_month_notdiscarded$dateusedforstatisticsiso)
meas_disaggregated_month_notdiscarded$isnotification <- as.factor(meas_disaggregated_month_notdiscarded$isnotification)

meas_disaggregated_last12months_notdiscarded$dateusedforstatisticsiso <- anydate(meas_disaggregated_last12months_notdiscarded$dateusedforstatisticsiso)
meas_disaggregated_last12months_notdiscarded$isnotification <- as.factor(meas_disaggregated_last12months_notdiscarded$isnotification)

meas_disaggregated_calendaryear_notdiscarded$dateusedforstatisticsiso <- anydate(meas_disaggregated_calendaryear_notdiscarded$dateusedforstatisticsiso)
meas_disaggregated_calendaryear_notdiscarded$isnotification <- as.factor(meas_disaggregated_calendaryear_notdiscarded$isnotification)

glimpse(meas_disaggregated_month_notdiscarded) 
```

3/ full_join()
```{r}
meas_combined_month_notdiscarded <- full_join(meas_month_notdiscarded, meas_disaggregated_month_notdiscarded)
meas_combined_last12months_notdiscarded <- full_join(meas_last12months_notdiscarded, meas_disaggregated_last12months_notdiscarded)
meas_combined_calendaryear_notdiscarded <- full_join(meas_calendaryear_notdiscarded, meas_disaggregated_calendaryear_notdiscarded) 
```

# Save working files

Discarded 
```{r} 
# month
csvFileName <- paste("../2 data/2 processed data/discarded/meas_month_discarded", yearofreport, monthofreport, ".csv",sep="_")  
write_csv(meas_month_discarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/discarded/meas_month_discarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_month_discarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/discarded/meas_agg_month_discarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_agg_month_discarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/discarded/meas_agg_month_discarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_agg_month_discarded,rdsFileName)

# last 12 months
csvFileName <- paste("../2 data/2 processed data/discarded/meas_last12months_discarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_last12months_discarded, csvFileName)
rdsFileName <- paste("../2 data/2 processed data/discarded/meas_last12months_discarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_last12months_discarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/discarded/meas_agg_last12months_discarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_agg_last12months_discarded, csvFileName)
rdsFileName <- paste("../2 data/2 processed data/discarded/meas_agg_last12months_discarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_agg_last12months_discarded,rdsFileName)

# calendaryear
csvFileName <- paste("../2 data/2 processed data/discarded/meas_calendaryear_discarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_month_discarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/discarded/meas_calendaryear_discarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_calendaryear_discarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/discarded/meas_agg_calendaryear_discarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_agg_month_discarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/discarded/meas_agg_calendaryear_discarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_agg_calendaryear_discarded,rdsFileName)
```

Not discarded 
```{r}
# month
csvFileName <- paste("../2 data/2 processed data/cases/meas_month_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_month_notdiscarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_month_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_month_notdiscarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/cases/meas_agg_month_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_agg_month_notdiscarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_agg_month_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_agg_month_notdiscarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/cases/meas_combined_month_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_combined_month_notdiscarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_combined_month_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_combined_month_notdiscarded,rdsFileName)


# last 12 months 
csvFileName <- paste("../2 data/2 processed data/cases/meas_last12months_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_last12months_notdiscarded, csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_last12months_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_last12months_notdiscarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/cases/meas_agg_last12months_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_agg_last12months_notdiscarded, csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_agg_last12months_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_agg_last12months_notdiscarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/cases/meas_combined_last12months_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_combined_last12months_notdiscarded, csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_combined_last12months_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_combined_last12months_notdiscarded,rdsFileName)

# calendaryear
csvFileName <- paste("../2 data/2 processed data/cases/meas_calendaryear_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_calendaryear_notdiscarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_calendaryear_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_calendaryear_notdiscarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/cases/meas_agg_calendaryear_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_agg_calendaryear_notdiscarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_agg_calendaryear_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_agg_calendaryear_notdiscarded,rdsFileName)

csvFileName <- paste("../2 data/2 processed data/cases/meas_combined_calendaryear_notdiscarded", yearofreport, monthofreport, ".csv",sep="_")
write_csv(meas_combined_calendaryear_notdiscarded,csvFileName)
rdsFileName <- paste("../2 data/2 processed data/cases/meas_combined_calendaryear_notdiscarded", yearofreport, monthofreport, ".rds",sep="_")
write_rds(meas_combined_calendaryear_notdiscarded,rdsFileName)
```
